Article Title

### 2014-01-26 plos.tex
Kreso's changes of the Discussion.

\section*{Discussion}
\paragraph{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often expressed as a graph of pairwise interactions. The intended purpose of many studies of functional connectivity is to estimate the circuit's anatomical connectivity solely from the observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections or shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are still ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from a larger subset of neurons in the circuit, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the rest of the network. Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.  Such conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, individual neurons' stimulus drive, and a network of synaptic couplings \cite{Pillow:2008}.  Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .  Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimal other assumptions about the distribution: the investigator need only choose which statistics to match.  Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of neurons.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the interaction between a pair of neurons on the activity of the other neurons under a given probabilistic model, the recording must include all the other neurons with which the pair can potentially interact. Otherwise, interactions with unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, the most successful applications of statistical models of population activity have been in \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations have been possible \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the conditional dependencies between pairs of cells. Perhaps for this reason, partial correlations have not been previously used to describe the functional connectivity between cortical neurons.  Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.

Correct inference of causal interactions between neurons depends on correct mathematical representations of the interactions in the statistical model and the completeness of the recorded population.  It is not yet clear which approach provides best correspondence to the anatomical connectivity in the circuit, with little experimental evidence yet available to address this question.  The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without a statement about its anatomical implications.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of the functional connectivity graphs vary with the method of their inference (Zalesky:2012); until the physiological meaning of the functional connectivity is established, the validity of such analyses remains in question.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with advanced spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of the inferred spike rates: better accuracy can be achieved when calcium signals are analyzed on the temporal scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may open new opportunities for more incisive measurements of functional connectivity than with electrophysiological recordings.

The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and maximum entropy models on the binary domain).  Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .  Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without a statement about underlying mechanisms.

In our study, we sought to infer pairwise functional connectivity networks among neurons in cortical microcircuits. We hypothesized that partial correlations would correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we a priori favored solutions with sparse partial correlations.  The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.  Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.

\paragraph{Functional connectivity as coactivations}
Another approach to describe the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}. Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}. Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \ie groups of tightly interconnected groups of cells involved in a common computation.  Coactivation patterns could also be explained as shared input from unobserved parts of the circuit.  Coactivations could also be explained by global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.

Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but analysis of coactivations method shifts the focus from detailed interactions to the collective behavior.

In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.

@@

In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.  The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.
\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.  The problem becomes increasingly worse as the size of the recorded population of neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As the recorded neuronal population sizes increased gradually, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.  However, with ambitious projects such as the BRAIN initiative currently underway \cite{Alivisatos:2013}, aiming to record from significantly larger populations, the problem must be addressed by regularizing the solution. Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data. However, limiting the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

@@

\paragraph{Model selection}
In our study, the covariance matrix estimators were evaluated with respect to the cross-validated normal log likelihood.  However, this does not limit the applicability of its conclusions to normal distributions. Indeed the major findings in this paper could be reproduced with respect to other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).  Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.  If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, a simple model can produce a better estimate even if it does not correctly represent the real dependencies in the data than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the inference that it reveals a better representation of dependencies in the data.  This caveat is related to the phenomenon known as \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

\paragraph{Physiological interpretation and future directions}
Although the graph of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred graph of interactions differed substantially from the graph of most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator revealed a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}).  These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.  The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.

The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.  Since the distinction led to improved estimation, the two types of components could be further investigated.  For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions. Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.  When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.  The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity. Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.

Here we showed that among several models we tested, we found that for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.  This finding valuable in itself: it can be used to improve studies aiming to understand the nature of population coding. Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.  For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry. Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2011,Ko:2013}  with our methods can be used to relate functional connectivity to the circuit architecture. Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation: this opens a window to describe what kinds of interactions are activated by stimulation conditions. Extending the estimation framework to temporal domain is important for future work.

Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred graph of interactions was substantially different from the graph of the most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).  These negative interactions are produced as explanations for low sample correlation values where high values are expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.  The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions.  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.

The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.   Since the distinction led to improved estimation, the two types of components could be further analyzed and related to physiological processes.  For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions. Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.  When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but has little effect on the pairwise interactions.  The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity. Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.

>>>>

\section*{Discussion}
\paragraph{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often represented as a graph of pairwise interactions. The goal of many studies of functional connectivity was to estimate  anatomical connectivity from  observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network. Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .  
%Although not informed by physiology, maximum entropy models are attractive because they account for the observed statistics with minimal other assumptions about the distribution: the investigator need only choose which statistics to match.  
Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

%Correct inference of causal interactions between neurons depends on correct mathematical representations of the interactions in the statistical model and the completeness of the recorded population.  
It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs vary with the method of their inference (Zalesky:2012). Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, the most statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations. 
% Despite this limitation, many studies have fitted probabilistic models to multielectrode recordings and analyzed the structure of the resulting conditional dependencies \cite{Tang:2008, Yu:2008}.  
\Kcomment{I thought this was a good way to end, and the last sentence went in a different direction.}

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.

The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and binary maximum entropy models).  Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .  Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without relating them to underlying mechanisms.

In our study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits. We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.  The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.  Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations. \Kcomment{I am not sure about the previous argument.}

\paragraph{Functional connectivity as coactivations}
Another approach to describing the functional connectivity of a circuit is to isolate patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}. Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}. Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \ie groups of tightly interconnected groups of cells involved in a common computation.  Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.

Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.  
In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.

@@

In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.  The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.  The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.  However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators. Regularization limits the solution to a smaller subspace to counteract the effect of the sampling noise in the empirical data. \Kcomment{This is not really true -- it doesn't limit the solution to a smaller subspace. It just shrinks it toward it.} However, limiting the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

@@

\paragraph{Model selection}
We evaluated the covariance matrix estimators using the cross-validated normal log likelihood.  However, this does not limit the applicability of its conclusions to normal distributions. Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:04} and Fig.~S\ref{supp:02}).  Other probabilistic models, fitted to the same data, would also serve as estimators of the covariance matrix.  \Kcomment{A probabilistic model is not an estimator.} If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, a wrong simple model can produce a better estimate than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.  This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

\paragraph{Physiological interpretation and future directions}
Although the graph of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model additional credence over other estimates of the functional connectivity.  The inferred graph of interactions differed substantially from the graph of most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator revealed a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}).  
%These negative interactions are often produced as explanations for low sample correlation values where high values were expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance. 
 The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions (\figref{fig:07}).  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether negative interactions preferentially involve inhibitory cells.

The  $C_{\sf sparse+latent}$ estimator decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.  Since the distinction led to improved estimation, the two types of components could be further investigated.  For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected in the common components/latent units but much less in pairwise interactions. Previous studies found that correlations measured under anesthesia retained little in common with correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.  When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but does not affect the pairwise interactions as strongly.  The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity. Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli.

Here we showed that among several models we tested, for cortical microcircuits, a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix.  This finding is valuable in itself: it can be used to improve studies aiming to understand the nature of population coding. Moreover, this estimation approach also provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit.  For example, we find a specific structure of negative interactions which may indicate inhibitory circuitry. Future experiment combining molecular markers for different cell types or follow-up multiple patching in slices \cite{Ko:2011,Ko:2013}  with our methods can be used to relate functional connectivity to the circuit architecture. Finally, the graphs of interactions derived \emph{in vivo} could potentially reveal how circuits reconfigure dynamically under different brain states or external stimulation. This opens a window to describe what kinds of interactions are activated by stimulation conditions. Extending the estimation framework to temporal domain is important for future work.

%Although the network of interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, resembles a network of synaptic interactions, it is unlikely that the anatomical and functional networks are identical.  The improved estimation performance gives the model extra credence over other estimates of the functional connectivity.  The inferred graph of interactions was substantially different from the graph of the most significant sample correlations \figref{05}{F, H, and I}.  The $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that are not present in the sample correlation matrix (\figref{05}{F}).  These negative interactions are produced as explanations for low sample correlation values where high values are expected based on common correlations with other neurons.  The sample correlation matrix does not draw attention to such surprisingly low correlations and potentially misses their physiological significance.  The resulting network had a larger fraction of negative interactions than in the sample correlation matrix. The negative interactions had a distinct spatial and functional organization from the positive interactions.  To determine potential physiological significance of the negative interactions, future experiments with labeled cell types could indicate whether the negative interactions emanate preferentially from inhibitory cells.
\Kcomment{The paragraph here (erased) was repeated from above.}

The estimator effectively decomposed the correlation structure into two distinct components: pairwise interactions and common coactivation modes denoted as \sq{latent units}.   Since the distinction led to improved estimation, the two types of components could be further analyzed and related to physiological processes.  For example, slow-wave network fluctuations due to anesthesia or slow-wave sleep could be strongly reflected on the common components/latent units but much less on the pairwise interactions. Previous studies found that correlations measured under anesthesia retained little in common to correlations measured in wakefulness and concluded that studies of the anesthetized brain did not inform about its behaviorally relevant functions \cite{Greenberg:2008}.  When conditioned on the brain state, optimal estimation of covariance matrices in dense cortical networks could reveal that slow-wave activity alters the common components but has little effect on the pairwise interactions.  The correlation structure also changes as a function of the stimulus condition \cite{Cotton:2013} on time scales that are much faster than changes in the anatomical connectivity. Descriptions of specific changes in the functional connectivity due to stimulus properties could shed light on how circuits orchestrate their activity to represent diverse stimuli. \Kcomment{This is also repeated.}
